
\section*{Acknowledgement}
\addcontentsline{toc}{section}{Acknowledgement}

Chen was supported in part by the project
``Bayesian Assessment of Safety Profiles for Pregnant Women From Animal
Study to Human Clinical Trial'' funded by U.S. Food and Drug Administration,
Office of Women's Health. The project was supported
in part by an appointment to the Research Participation Program at the
Center For Biologics Evaluation and Research administered by the Oak Ridge
Institute for Science and Education through an interagency agreement between
the U.S. Department of Energy and the U.S. Food and Drug Administration
\\

Chen was supported in part by
the Department of Ecology and Evolutionary Biology at the
University of Tennessee, Knoxville, and a grant from
the National Science Foundation (MCB-1120370.)
\\

Chen and Ostrouchov were supported in part by the project
``Visual Data Exploration and Analysis of Ultra-large Climate Data''
funded by U.S. DOE Office of Science
under Contract No. DE-AC05-00OR22725.
Ostrouchov, Schmidt, and Patel were supported in part by the project
``NICS Remote Data Analysis and Visualization Center''
funded by the Office of Cyberinfrastructure of the
U.S. National Science Foundation
under Award No. ARRA-NSF-OCI-0906324 for NICS-RDAV center.
\\

This work used resources of the Oak Ridge Leadership Computing Facility at the
Oak Ridge National Laboratory, which is supported by the Office of Science
of the U.S. Department of Energy under Contract No. DE-AC05-00OR22725.
This work also used resources of National Institute for Computational
Sciences at the University of Tennessee, Knoxville, which is supported
by the Office of Cyberinfrastructure of the U.S. National Science Foundation
under Award No. ARRA-NSF-OCI-0906324 for NICS-RDAV center.
This work used resources of the Newton HPC Program at the University of
Tennessee, Knoxville.
\\

We thank our colleagues from the Scientific Data Group, Computer
Science and Mathematics Division, Oak Ridge National Laboratory,
Hasan Abbasi, Jong Youl Choi, Scott Klasky, and Nobert
Podhorszki for discussing windows MPI systems, compiler issues,
dynamic libraries, and generally improving our knowledge of MPI
performance issues.
\\

We also thank Brian D. Ripley, Kurt Hornik, Uwe Ligges, and Simon Urbanek
from the \proglang{R} Core Team for discussing package release issues and
helping us solve portability problems on different platforms.
\\
